\Large \bf Cyclone: Warp Speed Replication for Key Value Stores

### 2017-07-21 paper/paper.tex
Updates to paper.

Cyclone is a high speed strongly consistent replicated write ahead logging
service specialized for key value stores such as Rocksdb. Cyclone provides the
order of magnitude better performance required to close the performance gap in
Figure~\ref{fig:problem} both on the storage and network side. Cyclone requires
\emph{a small amount of} directly attached Non Volatile Memory (NVM) on the host
- such as in the form of NVDIMMs~\cite{farm} or directly attached persistent
memory~\cite{pmfs}. Cyclone logs update operations to the key value store and
calls on the key value store to execute the update only when it has been
successfully replicated to majority quorum of replicas. All reads and write to
the key value store are always executed by the leader replica. A client side
stub provided by Cyclone automatically locates and directs read and write
requests to the leader replica, ensuring linearizable semantics from the key
value store. Using a replicated write ahead log to build a linearizable and
highly available key value store is not new [find some examples here]. The
novelty in Cyclone is that it accomplishes this without compromising on the
performance baseline of no logging or replication. Cyclone does this by
systematically addressing the sources of performance degradation in
Figure~\ref{fig:problem}.

@@

The network overheads for replicating a log entry to a quorum of machines stems
from the overhead of the TCP/IP stack in the linux kernel - itself the topic of
much research [dataplane OS stuff]. We therefore replase the TCP/IP stack with
DPDK, a userspace network library that allows direct IO of raw ethernet frames
to and from the network interface card (NIC) eliminating all overheads. We no
longer use TCP - since the rare network errors encountered in a typical
datacenter network are automatically handled by the replication protocol (RAFT)
as part of its failure model. Finally, we carefully design our implementation of
RAFT to relieve the CPU of all data copy operations, leaving the core to only
run the protocol and return an accept/reject decision for each received
packet. The set of optimizations to the network stack is decribed in
Section~\ref{sec:network}.

Both the storage (NVM, SSD) as well as the NIC expose significant parallelism
that is left unused when Cyclone is used to replicate a single log. To exploit
this parallelism we use Cyclone to replicate multiple logs, each using a
different instance of the RAFT consensus protocol. We show how to exploit the
commutativity inherent in key-value store interfaces to make use of this
parallelism in Cyclone's design in Section~\ref{parallelism}.

@@

\section{System Architecture}

\section{Storage}

\section{Network}

\section{Parallelism}

\section{Two Level Log}
\label{sec:2level}
A fundamental challenge in building an efficient replicated log is dealing with
network overheads. Cyclone's construction is predicated on the observation that
in the common case, protocols such as RAFT simply multicast the same entry
received from a client, from the leader replica to all follower replicas. This
suggests that an efficient way to implement a replicated log is to treat it as a
software packet switching problem. However peristing logs on flash represents an
impediment in that block storage devices are not good candidates to hold packets
being software switched. Our solution is to split the log into two levels. The
upper level is held in directly attached non-volatile memory (NVM)
such as NVDIMMs. The lower level is held on a flash SSD - we term the lower
level log as the flashlog. We periodically drain entries in batches from the NVM
log to the backing flashlog. An important consequence of this two level approach
is that we need only a limited amounts of NVM with the bulk of the log residing
on flash - an arrangement that maps well to the current relative costs of the
two types of memory. We begin by describing the NVM log and how we implement the
RAFT consensus protocol in the packet switching paradigm around it. We then
describe the flashlog and how we drain entries in batches from the NVM level to
the flash level.

@@

@@

>>>>

Cyclone is a high speed strongly consistent replicated write ahead logging
service specialized for key value stores such as Rocksdb. We describe Cyclone's
system architecture in Section~\ref{sec:sysarch}. Cyclone provides the order of
magnitude better performance required to close the performance gap in
Figure~\ref{fig:problem} both on the storage and network side. Cyclone requires
\emph{a small amount of} directly attached Non Volatile Memory (NVM) on the host
- such as in the form of NVDIMMs~\cite{farm} or directly attached persistent
memory~\cite{pmfs}. Cyclone logs update operations to the key value store and
calls on the key value store to execute the update only when it has been
successfully replicated to majority quorum of replicas. All reads and write to
the key value store are always executed by the leader replica. A client side
stub provided by Cyclone automatically locates and directs read and write
requests to the leader replica, ensuring linearizable semantics from the key
value store. Using a replicated write ahead log to build a linearizable and
highly available key value store is not new [find some examples here]. The
novelty in Cyclone is that it accomplishes this without compromising on the
performance baseline of no logging or replication. Cyclone does this by
systematically addressing the sources of performance degradation in
Figure~\ref{fig:problem}.

@@

The network overheads for replicating a log entry to a quorum of machines stems
from the overhead of the TCP/IP stack in the linux kernel - itself the topic of
much research [dataplane OS stuff]. We therefore replase the TCP/IP stack with
DPDK, a userspace network library that allows direct IO of raw ethernet frames
to and from the network interface card (NIC) eliminating all overheads. We no
longer use TCP - since the rare network errors encountered in a typical
datacenter network are automatically handled by the replication protocol (RAFT)
as part of its failure model. Finally, we carefully design our implementation of
RAFT to relieve the CPU of all data copy operations, leaving the CPU to only
run the protocol and return an accept/reject decision for each received
packet. The set of optimizations to the network stack is decribed in
Section~\ref{sec:network}.

Both the storage (NVM, SSD) as well as the NIC expose significant parallelism
that is left unused when Cyclone is used to replicate a single log. To exploit
this parallelism we use Cyclone to replicate multiple logs, each using a
different instance of the RAFT consensus protocol. We show how to exploit the
commutativity inherent in key-value store interfaces to make use of this
parallelism in Cyclone's design in Section~\ref{sec:parallelism}.

@@

\section{System Architecture}
\label{sec:sysarch}

\section{Storage}
\label{sec:storage}
A fundamental challenge in building an efficient replicated log is dealing with
network overheads. Cyclone's construction is predicated on the observation that
in the common case, protocols such as RAFT simply multicast the same entry
received from a client, from the leader replica to all follower replicas. This
suggests that an efficient way to implement a replicated log is to treat it as a
software packet switching problem. However peristing logs on flash represents an
impediment in that block storage devices are not good candidates to hold packets
being software switched. Our solution is to split the log into two levels. The
upper level is held in directly attached non-volatile memory (NVM)
such as NVDIMMs. The lower level is held on a flash SSD - we term the lower
level log as the flashlog. We periodically drain entries in batches from the NVM
log to the backing flashlog. An important consequence of this two level approach
is that we need only a limited amounts of NVM with the bulk of the log residing
on flash - an arrangement that maps well to the current relative costs of the
two types of memory. We begin by describing the NVM log and how we implement the
RAFT consensus protocol in the packet switching paradigm around it. We then
describe the flashlog and how we drain entries in batches from the NVM level to
the flash level.

@@

\section{Network}
\label{sec:network}

@@

\section{Parallelism}
\label{sec:parallelism}
